# 

# The daily COVID-19 cases data of Bolivia.

# 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(tidyverse)
library(coronavirus)
library(DT)
library(lubridate)
library(ggplot2)
library(plotly)
library(zoo)
data(coronavirus)
```

The following data table shows the number of new cases confirmed, recovered, dead every day in Bolivia. And it also contains the recover and death radio calculated from cumulative cases. 

You can try to type a date you are interested in and see the data of that day.

```{r interactive-table}
covid_bolivia <- coronavirus %>% 
  filter(country == "Bolivia") %>% 
  select(date, type, cases)

covid_bolivia_wide <- covid_bolivia %>% 
  pivot_wider(names_from = type, values_from = cases) 

ratios_bolivia <- covid_bolivia_wide %>% 
  filter(date > as.Date("2020-03-10")) %>% 
  mutate(recover_cum = cumsum(recovered), 
         confirm_cum = cumsum(confirmed), 
         death_cum = cumsum(death))

ratios_bolivia <- ratios_bolivia %>% 
  mutate(recover_ratio = round(recover_cum / confirm_cum, 2), 
         death_ratio = round(death_cum / confirm_cum, 2))

ratios_bolivia %>% 
  select(date, confirmed, confirm_cum, recovered, recover_ratio, 
         death, death_ratio) %>% 
  arrange(desc(date)) %>% 
  datatable(extensions = 'Responsive', 
            caption = "Daily COVID-19 Cases in Bolivia")
```

```{r ratioB}
ratios_bolivia <- ratios_bolivia %>% 
  mutate(recover_ratio = recover_cum / confirm_cum, 
         death_ratio = death_cum / confirm_cum)
```

```{r trend-plotB, fig.align='center'}
covid_bolivia <- covid_bolivia %>% 
  filter(date > as.Date("2020-03-10")) %>% 
  group_by(factor(type)) %>% 
  mutate(cases_cum = cumsum(cases)) %>% 
  ungroup()

plot_trend_bolivia <- covid_bolivia %>% 
  ggplot(aes(x = date, y = cases_cum / 1000, color = `factor(type)`)) + 
  geom_point(size = 0.5) + 
  scale_x_date(date_breaks = "1 month", date_labels = "%b") + 
  scale_y_continuous(breaks = c(20, 40, 60, 80, 100, 120)) + 
  ylab("cumulate cases (thousand)")


ggplotly(plot_trend_bolivia, tooltip = c("x", "y"))
```

# Data Story 1: Check the trends of newly confirmed cases by 7-day / 14-day moving average.

```{r 7day-avg}
covid_bolivia <- covid_bolivia %>% 
  filter(type == "confirmed") %>% 
  dplyr::mutate(seven_day_Avg = rollmean(cases, k = 7, fill = NA), 
                fourteen_day_Avg = rollmean(cases, k = 14, fill = NA)) 

plot_avg_bolivia <- covid_bolivia %>% 
  filter(date > as.Date("2020-03-10")) %>% 
  ggplot() + 
  geom_line(aes(x = date, y = seven_day_Avg), color = "blue") + 
  geom_line(aes(x = date, y = fourteen_day_Avg), color = "green") 

ggplotly(plot_avg_bolivia)
```


# Data story 2: Recovery / death ratio comparions between Chile and Bolivia.

```{r compare-recovery}
compare <- covid_chile_wide %>% 
  left_join(covid_bolivia_wide, by = 'date') %>% 
  filter(date > as.Date("2020-03-10")) %>% 
  mutate(recover_ratio_chile = cumsum(recovered.x) / cumsum(confirmed.x), 
         recover_ratio_bolivia = cumsum(recovered.y) / cumsum(confirmed.y))

compare_long <- compare %>% 
  select(date, recover_ratio_chile, recover_ratio_bolivia) %>% 
  rename(Chile = recover_ratio_chile, 
         Bolivia = recover_ratio_bolivia) %>% 
  pivot_longer(2:3, names_to = "Country", values_to = "recovery_ratio")

compare_long <- compare_long %>% 
  mutate(Week = week(date)) %>% 
  group_by(Week, Country) %>% 
  summarise(weekly_recovery_ratio = mean(recovery_ratio)) 

plot_compare <- compare_long %>% 
  ggplot(aes(x = Week, y = weekly_recovery_ratio, fill = Country)) + 
  geom_col(position = "dodge") + 
  scale_x_continuous(breaks = c(15, 20, 25, 30, 35, 40))

ggplotly(plot_compare)
  
```


```{r compare-death}
compare_death <- covid_chile_wide %>% 
  left_join(covid_bolivia_wide, by = 'date') %>% 
  filter(date > as.Date("2020-03-10")) %>% 
  mutate(death_ratio_chile = cumsum(death.x) / cumsum(confirmed.x), 
         death_ratio_bolivia = cumsum(death.y) / cumsum(confirmed.y))

compare_death_long <- compare_death %>% 
  select(date, death_ratio_chile, death_ratio_bolivia) %>% 
  rename(Chile = death_ratio_chile, 
         Bolivia = death_ratio_bolivia) %>% 
  pivot_longer(2:3, names_to = "Country", values_to = "death_ratio")

compare_death_long <- compare_death_long %>% 
  mutate(Week = week(date)) %>% 
  group_by(Week, Country) %>% 
  summarise(weekly_death_ratio = mean(death_ratio)) 

plot_compare_death <- compare_death_long %>% 
  ggplot(aes(x = Week, y = weekly_death_ratio, fill = Country)) + 
  geom_col(position = "dodge") + 
  geom_hline(yintercept = 0.03, color = "orange", linetype = "dashed") + 
  geom_hline(yintercept = 0.04, color = "orange", linetype = "dashed") + 
  scale_x_continuous(breaks = c(15, 20, 25, 30, 35, 40)) 

ggplotly(plot_compare_death)
```





